### 2020-10-27 report.Rmd
Cleaned up rmd

---
title: "Report on Effect Size Translation"
ipsum_meta:
  twitter_card: "Summary info for the Twitter Card"
  twitter_site: "\\@sitehandle"
  twitter_creator: "\\@creatorhandle"
  og_url: "https\\://example.com/open/graph/finalURLfor/this"
  og_description: "A modest size description of the content"
  og_image: "https\\://example.com/open/graph/imageURLfor/this"
output: 
  hrbrthemes::ipsum:
    toc: true
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.retina=2, echo=FALSE,message = FALSE,
                      warnings=FALSE, error = FALSE)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
library(hrbrthemes)
library(tidyverse)
library(targets)
library(gt)

update_geom_font_defaults(font_rc)
```

# Assumptions made

I made the following assumptions which may or may not be reasonable:

1.  That $\beta$ is equivilent to $r$ for $\beta < |.5|$ and undefined otherwise (never the case here)
2.  That, given the large sample sizes involved *Hedge's g* and *Cohen's d* where essentially equivalent and that sample sizes used to make up these effects were approximately equal in treatment and control groups.
3.  In the absence of information that could be used to construct a 2x2 contingency table that ORs and Relative Risk Ratios were equivalent and thus Relative Risk Ratios were converted to $r$ as if they were Odds ratios.
4.  External to any other information that all forms of SMD, weighted means etc. were equivilent to Cohen's d.

# Correspondence

```{r}
targets::tar_load(converted_data)

converted_data %>%
  group_by(std_eff_name, es) %>%
  count(sort = TRUE) %>%
  gt() %>%
  tab_header(
    title = "Coversion of effect sizes",
    subtitle = "New Effect Name - Old Effect Name"
  )
```

# Effect Size Summary: Point Estimates

```{r}

converted_data %>%
  group_by(std_eff_name) %>%
  summarize(Mean = mean(value_consensus_r, na.rm = TRUE),
            SD = sd(value_consensus_r, na.rm = TRUE),
            Min = min(value_consensus_r, na.rm = TRUE),
            Max = max(value_consensus_r, na.rm = TRUE)) %>%
  gt() %>%
  tab_header(
    title = "Effect Size Summary",
    subtitle = "Descriptives by Effect Size Type"
  ) %>%
  fmt_number(columns = 2:5,
             decimals = 3)

```

# Project Structure
```{r}
targets::tar_visnetwork()
```