### 2022-01-07 reports/manuscript_distill.Rmd

---
title: "Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review"
author:
- name: Taren Sanders
  affiliation: Institute for Positive Psychology and Education, ACU
  affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
  orcid_id: 0000-0002-4504-6008
- name: Michael Noetel
  affiliation: School of Health and Behavioural Sciences, ACU
  affiliation_url: https://www.acu.edu.au/about-acu/faculties-directorates-and-staff/faculty-of-health-sciences/school-of-behavioural-and-health-sciences
  orcid_id: 0000-0002-6563-8203
- name: Philip Parker
  affiliation: Institute for Positive Psychology and Education, ACU
  affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
  orcid_id: 0000-0002-4604-8566
- name: Borja Del Pozo Cruz
- name: Stuart Biddle
- name: Rimante Ronto
- name: Ryan Hulteen
- name: Rhiannon Parker
- name: George Thomas
- name: Katrien De Cocker
- name: Jo Salmon
- name: Kylie Hesketh
- name: Nicole Weeks
- name: Hugh Arnott
- name: Emma Devine
- name: Roberta Vasconcellos
- name: Chris Lonsdale
  affiliation: Institute for Positive Psychology and Education, ACU
  affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
  orcid_id: 000-0002-2523-5565
date: '`r format(Sys.Date())`'
repository_url: https://github.com/Motivation-and-Behaviour/screen_umbrella
bibliography: combined.bib
csl: vancouver-brackets.csl
nocite: "`r params$nocite_list`"
params:
  render_img: true
  nocite_list : true
  inc_tabs: true
---

```{r setup, include=FALSE}
# This allows distill to render from a targets pipeline
Sys.setenv("RSTUDIO_VERSION" = '1.4.1725')

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

options(dplyr.summarise.inform = FALSE)

# Uncomment to run outside of pipeline
# library(tidyverse)
# library(targets)
# library(xfun)
```

```{r child = 'manuscript_content.Rmd'}
```

# Appendix

#### Figures

* [Prisma Diagram](https://drive.google.com/file/d/1ezgeAr2WPhy6MNSESS8bhg3VMOrLAUC7/view?usp=sharing)
* [Education Outcomes](https://drive.google.com/file/d/1ODsKf3jjEGNCpUTz5tju8AHRpk_xIz-e/view?usp=sharing)
* [Health-Related Outcomes](https://drive.google.com/file/d/1wdKzliWLEFuVveAj6-MzOAsoqBhnkRJW/view?usp=sharing)

#### Tables

* [Descriptive Table](https://drive.google.com/file/d/1t5XvwY2kzsOsbZ2FxbGPsYEecZz5PgD1/view?usp=sharing)
* Supplemental tables (TBC)

#### Supplementary Files

* [Supplementary File 1: List of Reviews Without Meta-analyses](https://drive.google.com/file/d/1JhMsySEyfby5MtcUcIVpp0LjDDfTA5IH/view?usp=sharing)
* [Supplementary File 2: List of Exposures](https://drive.google.com/file/d/12rB0ihW9Hi1kxM-8AIBfjdCMO6TqP28n/view?usp=sharing)
* [Supplementary File 3: Education Outcomes Not Meeting Statistical Certainty](https://drive.google.com/file/d/1wUotZRh-jvdBJh5dkstkMtvcz7DPn-vS/view?usp=sharing)
* [Supplementary File 4: Health Outcomes Not Meeting Statistical Certainty](https://drive.google.com/file/d/1-3frrZ3woWjPqkO84Dn3_1CCaE8thIZn/view?usp=sharing)

#### Links

* [Manuscript Google Doc](https://docs.google.com/document/d/1NN4yJWQVepNjbX0mND4aidWWXcCDkg-eMepRnWdtQ9o/edit#)
* [Online Report](https://motivation-and-behaviour.github.io/screen_umbrella/)

# References