### 2022-01-07 reports/manuscript_papaja.Rmd

---
title             : "Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review"
shorttitle        : "Benefits and risks of electronic screens"

author: 
  - name          : "Taren Sanders"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "33 Berry St, North Sydney, NSW, Australia"
    email         : "Taren.Sanders@acu.edu.au"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data curation
      - Writing - Original Draft
  - name          : "Michael Noetel"
    affiliation   : "2"
    role:
      - Conceptualization
      - Writing - Original Draft
  - name          : "Philip Parker"
    affiliation   : "1"
    role:
      - Conceptualization
      - Writing - Original Draft
  - name          : "Borja Del Pozo Cruz"
    affiliation   : "3"
    role:
      - Writing - Review & Editing
  - name          : "Stuart Biddle"
    affiliation   : "4"
    role:
      - Writing - Review & Editing
  - name          : "Rimante Ronto"
    affiliation   : "5"
    role:
      - Writing - Review & Editing
  - name          : "Ryan Hulteen"
    affiliation   : "6"
    role:
      - Writing - Review & Editing
  - name          : "Rhiannon Parker"
    affiliation   : "7"
    role:
      - Writing - Review & Editing
  - name          : "George Thomas"
    affiliation   : "8"
    role:
      - Writing - Review & Editing
  - name          : "Katrien De Cocker"
    affiliation   : "5"
    role:
      - Writing - Review & Editing
  - name          : "Jo Salmon"
    affiliation   : "9"
    role:
      - Writing - Review & Editing
  - name          : "Kylie Hesketh"
    affiliation   : "9"
    role:
      - Writing - Review & Editing
  - name          : "Nicole Weeks"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
  - name          : "Hugh Arnott"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
  - name          : "Emma Devine"
    affiliation   : "10"
    role:
      - Writing - Review & Editing
  - name          : "Roberta Vasconcellos"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
  - name          : "Chris Lonsdale"
    affiliation   : "1"
    role:
      - Writing - Original Draft
      - Project Administration

affiliation:
  - id            : "1"
    institution   : "Institute for Positive Psychology and Education, Australian Catholic University"
  - id            : "2"
    institution   : "School of Health and Behavioural Sciences, Australian Catholic University"
  - id            : "3"
    institution   : "Department of Sport Science and Clinical Biomechanics, University of Southern Denmark"
  - id            : "4"
    institution   : "School of Psychology and Counselling, University of Southern Queensland"
  - id            : "5"
    institution   : "Department of Health Systems and Populations, Faculty of Medicine, Health and Human Sciences, Macquarie University"
  - id            : "6"
    institution   : "School of Kinesiology, Louisiana State University"
  - id            : "7"
    institution   : "School of Medicine and Health, Sydney University"
  - id            : "8"
    institution   : "Faculty of Health Sciences, Curtin University"
  - id            : "9"
    institution   : "Institute for Physical Activity and Nutrition, Deakin University"
  - id            : "10"
    institution   : "The Matilda Centre for Research in Mental Health and Substance Use, University of Sydney"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["combined.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : yes

documentclass     : "apa6"
classoption       : "man"
csl               : vancouver-brackets-meta.csl
output            : papaja::apa6_pdf
nocite            : "`r params$nocite_list`"
params:
  render_img: true
  nocite_list : true
  inc_tabs: false
---

```{r setup, include=FALSE}
Sys.setenv("RSTUDIO_VERSION" = '1.4.1725')
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

options(dplyr.summarise.inform = FALSE)

# Uncomment to run outside of pipeline
# library(tidyverse)
# library(targets)
# library(xfun)
```

```{r child = 'manuscript_content.Rmd'}
```

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
References marked with an asterisk indicate studies which were included in the synthesis.
<div id = "refs"></div>
\endgroup