### 2023-03-13 README.Rmd
Revisions for Nature Human Behaviour (#27)

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review

<!-- badges: start -->
<!-- badges: end -->

Contains all of the project files to conduct the analysis and generate the manuscript for the umbrella review.

## Using this Repo

1. Clone the repository, ensuring that you are using the latest version.
1. Open directory and run `renv::restore()` to download/update the required packages.
1. To build the project, run `targets::tar_make()`.

## Links

* [PROSPERO Registration](https://www.crd.york.ac.uk/prospero/display_record.php?RecordID=76051)
* [Open Science Framework](https://osf.io/3ubqp/)